*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-01 Code/Manuscript_O2.Rmd
some edits

Current niches of *P. marinus* strains span ocean water columns [@veldhuisVerticalDistributionPigment1990; @chisholmNovelFreelivingProchlorophyte1988; @partenskyProchlorococcusMarinePhotosynthetic1999] and extend into regions with low dissolved oxygen concentrations [@partenskyProchlorococcusAdvantagesLimits2010; @goerickeNovelNicheProchlorococcus2000; @lavinNovelLineagesProchlorococcus2010, @ulloaCyanobacteriumProchlorococcusHas2021; @garcia-robledoCrypticOxygenCycling2017].

@@

The OceansMap Protein Portal is an open access online data repository (Woods Hole Oceanographic Institute, WHOI) of mass spectroscopy data on marine microbial peptides, sampled from various depths and locations worldwide [@saitoDevelopmentOceanProtein2021]. We screened a subset of the OPP for proteins annotated as from *Prochlorococcus* strains, to identify differential strategies employed by strains living at varying depths and oxygen levels within the marine water column.  We focused on proteins mediating photosynthesis and protein metabolism. The samples for metaproteomic analyses were collected from 12 locations in the tropical North Pacific ocean along 150 W from 18 N of the equator between October 1, 2011 and October 25, 2011 during the voyage of the R/V Kilo Moana MetZyme cruise KM1128 (https://www.rvdata.us/search/cruise/KM1128; original datasets in the Biological and Chemical Oceanography Data Management Office repository; https://www.bco-dmo.org/project/2236).

Collection and treatment of protein samples were performed by personnel at the WHOI according to the protocols described by Saito *et al*. [@saitoMultipleNutrientStresses2014; @saitoNeedlesBlueSea2015]. Briefly, samples of seawater from depths of 20, 40, 50, 60, 70, 80, 90, 120, 150, 200, 250, 300, 380, 400, 500, 550, 600, and 800 m below the ocean surface were pumped through a 0.2 µm filter, preserved in RNAlater and frozen at -80°C until extraction. Proteins were extracted from the filter in an SDS-based detergent, embedded in tube gel, alkylated and reduced prior to in-gel trypsin digestion. Peptide spectra were generated using a Q-Exactive Orbitrap Mass Spectrometer, searched in the SEQUEST CITATIONXXXX and labelled with the most likely protein and species annotation from Uniprot.  Oxygen levels at the location of sampling were recorded.

@@

## Quantitative protein immunoblotting {.unnumbered}

Quantitative protein immunoblots were used to determine the concentrations of 5 target proteins; FtsH, PsbA, PetC, PsaC and RbcL in *P. marinus* MED4 and MIT9313 cultures grown in the PSI Multicultivator under imposed PAR light levels of 30 and 100 µmol photons m^-2^ s^-1^, spectral wavebands of 660 and 450 nm and [O~2~] of 2.5 and 250 µM. 40 mL of culture were transferred into a 50 mL falcon tube and 40 µL of 10 % Pluronic F68 acid was added to the cell suspension. The cell suspension was centrifuged for 30 minutes at 17664 g using a J-26XP centrifuge (Beckman Coulter, Brea, CA, USA) equipped with a J-10 rotor. The supernatant was removed and the resulting pellet was flash frozen in liquid nitrogen and stored at -80 °C for later use.

Protein samples were thawed and placed in 2 mL FastPrep®-24 bead lysing matrix D vials (MP Biomedicals, Irvine, CA, USA; SKU 116913050) with 200 µL of 1X protein extraction buffer (0.0281 mol L^−1^ TRIS base buffer, 0.0211 mol L^−1^ TRIS HCl buffer, 0.0735 mol L^−1^ lithium dodecyl sulfate, 1.08 mol L^−1^ glycerol, 0.5 mmol L^−1^ ethylenediaminetetraacetic acid, 0.1 mg·mL−1 PefaBloc SC (AEBSF) protease inhibitor) modified from Brown *et al*. [@brownFluxCapacitiesAcclimation2008]. Cells were lysed for 3 cycles of 60 seconds at 6.5m s-1 with 60 second intervals on ice between cycles.  The lysing vials were centrifuged at 16100 x *g* using a tabletop Mikro 20 centrifuge (Andreas Hettich GmbH & Co., Tuttlingen, Germany) for 5 minutes, the supernatant was transferred to a 600 µL microcentrifuge tube and centrifuged again at 16100 x *g* for 3 minutes. 40 µL of the supernatant were subsampled into labeled vials and stored at -80 °C until analysis. The bicinchoninic acid (BCA) assay was used for total protein determination using a calibration curve of bovine gamma globulin (BGG; Bio-Rad, Hercules, CA, USA) standard ranging from 0 to 1.0 mg mL^-1^ range.

All samples and appropriate standards were prepared as described in [@brownFluxCapacitiesAcclimation2008]. Standards were obtained from Agrisera (Vännäs, Sweden; FtsH AS11 1789S; PsbA AS01 016S; PetC AS08 330S; PsaC AS04 042S; RbcL AS01 017S).  Between 0.1 to 4.0 µg of total protein was loaded onto 15 well Invitrogen Bolt 4-12% acrylamide Bis-Tris gels (ThermoFisher Scientific, Waltham, MA, USA). All gels were run for 25 minutes at 200V using Invitrogen Bolt Mini Gel tanks and Invitrogen Bolt MES SDS running buffer (ThermoFisher Scientific, Waltham, MA, USA) and transferred for 60 minutes at 20V onto a PVDF membrane (Bio-Rad, Hercules, CA, USA; SKU 1620177). Blots were blocked overnight in 2% ECL blocking solution as described by @brownFluxCapacitiesAcclimation2008. The following morning, each blot was incubated in their appropriate antibody obtained from Agrisera (Vännäs, Sweden; FtsH  1:5000 dilution, AS11 1789; PsbA  1:10000 dilution, AS05 084; PetC 1:10000 dilution, AS08 330; PsaC 1:5000 dilution, AS10 939; RbcL 1:10000 dilution, AS03 037) and a 1:15000 dilution of the secondary antibody, goat anti-rabbit HRP conjugated (Bio-Rad, Hercules, CA, USA; SKU 1706515, lot 64230791).  Blots were incubated in Amersham ECL Select chemiluminescent substrate (Danaher, Washington, DC, USA) and imaged in a Bio-Rad VersaDoc imager (Bio-Rad, Hercules, CA, USA). Images were quantitated using the ImageJ2 software (version 1.53c).

@@

XXXMore TextXXX from Amanda & AuroraXXX
Proteins derived from Clade HLXXX ecotypes of *P. marinus* were detected in OMZ at depths up to 200 meters,  with O~2~ of 15 µM. The extent to which [O~2~] defines the niches occupied by different *P. marinus* ecotypes, as compared to potentially covarying environmental variables like photoperiod, light spectrum, and light level, is poorly described.

@@

NATLProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (Clade HLI), SS120 (Clade LLIII) and MIT9313 (Clade LLIV).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

### MED4 Clade HLI {.unnumbered}

Under 250 µM O~2~ the growth rates of *P. marinus* MED4, Clade HLI, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), across all spectral wavebands. No growth was observed under any imposed conditions under a 4 h photoperiod. The maximum growth rate (µ~max~) was `r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 180 µE blue light and 16 h photoperiod.

@@

The GAM model in Figure \@ref(fig:MED4GAM) summarizes MED4 growth responses to red (A,B) or blue (C,D) peak PAR and photoperiod across 2 imposed oxygen concentrations.  Under 250 µM O~2~ MED4 achieved fastest growth rates above peak blue light of ~180 µmol photons m^-2^ s^-1^, and the longest photoperiod of 16, indicated by the 0.64 d^-1^ contour line representing the 90^th^ percentile of maximum achieved growth rate (Figure \@ref(fig:MED4GAM)C).  Growth decreased with decreasing photoperiod and decreasing peak PAR. Under red light growth was generally slower but the pattern of growth responses to photoperiod and PAR was similar (Figure \@ref(fig:MED4GAM)A).  Note the exclusion of MED4 from growth under 4 h photoperiod under both red and blue light (Figure \@ref(fig:MED4GAM)). Under 25 µM O~2~ MED4 showed similar growth responses, but was excluded from both 4 and 8 h photoperiods.  MED4 did not grow under 2.5 µM O~2~, so no GAM model was run. 
Considering the range of PAR levels, and spectral bands that MED4 can utilize, MED4 can inhabit not just shallow depths, where light levels are high, but also deeper regions, characterized by a lower level of blue light, subject to the limitation of a photoperiod of more than 4 h, even after depth attenuation of light. The photoregimes of winter temperate zones, due to shorter photoperiods, exclude MED4 from growth at any depth, however temperate photoperiods and light levels for the remainder of the year are potentially adequate to support MED4 growth, if water temperatures warm into the Clade HLI tolerance range.

@@

### SS120 Clade LLIII {.unnumbered}

Under 250 µM O~2~, growth rates of *P. marinus* SS120, Clade LLIII, increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and across all spectral wavebands (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^ but showed growth inhibition at 16 h red light photoperiod.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white LED, red or blue light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

@@

### MIT9313 Clade LLIV {.unnumbered}

Under 250 µM O~2~ growth rates of *P. marinus* MIT9313, Clade LLIV, increased with longer
photoperiods, under low 30 µmol photons m^-2^ s^-1^ peak PAR, (Figure \@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiods. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while MIT9313 showed only marginal growth under white LED and red light at 180 µmol photons m^-2^ s^-1^ peak PAR, under the 8 h photoperiod, consistent with Moore *et al*. [@moorePhotophysiologyMarineCyanobacterium1999]. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 30 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod.

@@

### MED4 Clade HLI {.unnumbered}

@@

### SS120 Clade LLIII {.unnumbered}

@@

### MIT9313 Clade LLIV {.unnumbered}

@@

Under full atmospheric [O~2~] and blue light, LL clades of *Prochlorococcus* are restricted to growth under low light, in part because they suffer photoinhibition of Photosystem II (PSII) through several paths, including direct absorbance of UV or blue light, in parallel with generation of Reactive Oxygen Species (ROS) if the electron flow is slowed [@aroPhotoinhibitionPhotosystemII1993], producing damaging singlet oxygen (^1^O~2~) [@aroPhotoinhibitionPhotosystemII1993; @soitamoPhotoinhibitionMarinePicocyanobacteria2017; @murphyPhotoinactivationPhotosystemII2017; @hakalaEvidenceRoleOxygenevolving2005]. Repair of photoinactivated PSII relies on the removal of damaged PsbA [@mannInvolvementFtsHHomologue2000; @adamFtsHProteasesChloroplasts2005], followed by reassembly with newly synthesized PsbA [@nixonRecentAdvancesUnderstanding2010]. Degradation of PsbA is a rate-limiting step in recovery from photoinhibition [@kanervoD1ProteinDegradation1993], mediated largely by a heterohexamer of (FtsH12)~3~, a membrane-bound (Sacharz et al., 2015)(Zak et al., 2001) metalloprotease  (Chiba et al., 2002) (Yoshioka-Nishimura and Yamamoto, 2014).  [@adamFtsHProteasesChloroplasts2005; @boehmSubunitOrganizationSynechocystis2012a; @nixonRecentAdvancesUnderstanding2010; @komendaExposedNTerminalTail2007] (Pisareva et al., 2007; Sacharz et al., 2015),  @mannInvolvementFtsHHomologue2000, (AMANDA REF from ALGATECH).

*Prochlorococcus* genomes encode 4 FtsH proteins, henceforth referred to as FtsH1-4, homologs to the characterized FtsH isoforms of the model freshwater cyanobacterium *Synechocystis* FtsH, and with presumably parallel functions (Table \@ref(tab:FtsHTable)). Upon a shift to higher light HLI MED4 upregulated expression of FtsH1 and FtsH2 [@bonisteelStrainSpecificDifferences2018], homologs to the *Synechocystis* slr0228 and slr1604, implicated in PSII repair. In contrast, representative LLIV strain MIT9313 showed less overall expression of the FtsH proteases, and thus has fewer FtsH serving each photosystem.  Furthermore, MIT9313 expressed primarily FtsH3, homologous to *Synechocystis* slr1463 involved in PSI biogenesis, and  FtsH expression did not increase in response to light stress in MIT9313. Through adaptation to steady low light, Clade LLIV *Prochlorococcus* instead allocate resources to processes other than dynamic regulation of PSII repair.

@@

Ocean detections of proteins mediating protein metabolism support this interpretation of distinct FtsH function across clades of *P. marinus*. Ribosome proteins from Clade HLI MED4, Clade LLIII SS120 and Clade LLIV MIT9313 show generally similar patterns vs. [O~2~] and depth, a proxy for peak PAR.  FtsH3, inferred to mediate PSI assembly, likewise shows a similar pattern between MED4 and MIT9313.  But only MED4 shows detected presence of the FtsH1 & FtsH2 isoforms inferred to mediate PSII repair, and then only in near-surface samples subject to higher light.  Furthermore, even though MIT9313 grows (Figure \@ref(fig:MIT9313GAM)), and is detected in the ocean at low [O~2~] (Figure \@ref(fig:NATLProchloroPhotosynthDepthO2)), no FtsH from MIT9313 is detected at low [O~2~], suggesting limited requirement for protein turnover under low [O~2~].

@@

NATLProchloroProteinMetabDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* protein metabolism complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (Clade HLI), SS120 (Clade LLIII) and MIT9313 (Clade LLIV).",
                       "Columns show detections of proteins annotated as FtsH Protease Complexes (FtsH1, FtsH2, FtsH3) or the Ribosome).",
                       "Culture growth experimental conditions indicated by horizontal grey lines for depths approximating Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^) and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

XXXX Mireille to finish Fig. caption when figure is complete. Add colour ? Remove []? XXX

```{r  MED4ProteinAnnotated, fig.cap = MED4ProteinAnnotated_cap, fig.height = 6, fig.width = 8, echo = FALSE}

MED4ProteinAnnotated_cap <- glue("**Bar graph indicating protein concentrations determined from lab grown *Prochlorococcus marinus* MED4 (High-Light (HLI) near surface clade.  **",
                                    "Growth Photosynthetically Active Radiation (PAR) (µmol photons m^-2^ s^-1^) and spectral wavelength are in rows;  2 levels of imposed growth dissolved O~2~ concentrations (µM) are in columns.",
                                    "Numbers over each bar are fmole target protein per ug total protein.",
                                    .sep = " ")

knitr::include_graphics(file.path(FiguresIn, "MED4ProteinAnnotated.png"))

XXXX Mireille to finish Fig. caption when figure is complete. Add colour ? Remove []? XXX

```{r  MIT9313ProteinAnnotated, fig.cap = MIT9313ProteinAnnotated_cap, fig.height = 6, fig.width = 8, echo = FALSE}

MIT9313ProteinAnnotated_cap <- glue("**Bar graph indicating protein concentrations determined from lab grown *Prochlorococcus marinus* MIT9313 (Low-Light (LLIV) deep ocean clade).  **",
                                    "Growth Photosynthetically Active Radiation (PAR) (µmol photons m^-2^ s^-1^) and spectral wavelength are in rows;  2 levels of imposed growth dissolved O~2~ concentrations (µM) are in columns.",
                                    "Numbers over each bar are fmole target protein per ug total protein.",
                                    .sep = " ")

knitr::include_graphics(file.path(FiguresIn, "MIT9313ProteinAnnotated.png"))

@@

*Prochlorococcus marinus* SS120, a LLIII clade representative, showed an interactive inhibition of growth by oxygen and cumulative diel PUR, with a higher tolerance for higher cumulative diel PUR under 25 µM O~2~, compared to 250 µM O~2~. Thus, SS120 can exploit habitats with O~2~ levels spanning 25 µM O~2~ to 250 µM O~2~, including high PAR environments within OMZ.  SS120 is likely excluded from higher [O~2~] and higher PAR because of genomic limitations on capacity for DNA repair (Figure \@ref(fig:DNARepairFig)), and possibly by limited capacity for synthesis of reactive oxygen quenchers (Figure  \@ref(fig:ProchlorococcusEnzymeKms)).  Our growth results are supported by @lavinNovelLineagesProchlorococcus2010 who found evidence of LLII/III
and LLIV ecotypes, using terminal restriction fragment length polymorphism analyses, at depths above 40 m, where light levels are
higher, within OMZ, and by Figure \@ref(fig:NATLProchloroPhotosynthDepthO2) showing PSII protein subunits annotated as derived from SS120 at all depths ranging
from 20 to 200 m and all [O~2~] in an OMZ of the tropical North Pacific Ocean. SS120, grew under photoperiods longer than 4 h and showed increasing growth rate with increasing photoperiods. Therefore, SS120 can potentially exploit the longer summer photoperiods in temperate zones but again cannot sustain growth in winter through its requirement for more than 4 h of light per day, even if water temperature warms into the Clade LLIII tolerance range.

>>>>

Current niches of *P. marinus* strains span ocean water columns [@veldhuisVerticalDistributionPigment1990; @chisholmNovelFreelivingProchlorophyte1988; @partenskyProchlorococcusMarinePhotosynthetic1999] and extend into regions with low dissolved oxygen concentrations [@partenskyProchlorococcusAdvantagesLimits2010; @goerickeNovelNicheProchlorococcus2000; @lavinNovelLineagesProchlorococcus2010; @ulloaCyanobacteriumProchlorococcusHas2021; @garcia-robledoCrypticOxygenCycling2017].

@@

The OceansMap Protein Portal is an open access online data repository (Woods Hole Oceanographic Institute, WHOI) of mass spectroscopy data on marine microbial peptides, sampled from various depths and locations worldwide [@saitoDevelopmentOceanProtein2021]. We screened a subset of the OPP for proteins annotated as from *Prochlorococcus* strains, to identify differential strategies employed by strains living at varying depths and oxygen levels within the marine water column.  We focused on proteins mediating photosynthesis and protein metabolism from depths of 20 to 200 m below the ocean surface. The samples for metaproteomic analyses were collected from 12 locations in the tropical North Pacific ocean along 150 W from 18 N of the equator between October 1, 2011 and October 25, 2011 during the voyage of the R/V Kilo Moana MetZyme cruise KM1128 (https://www.rvdata.us/search/cruise/KM1128; original datasets in the Biological and Chemical Oceanography Data Management Office repository; https://www.bco-dmo.org/project/2236). Oxygen concentration levels at the location of sampling were recorded. The methodology for sample collection and peptide analysis are described by Saito *et al*. [@saitoMultipleNutrientStresses2014; @saitoNeedlesBlueSea2015].

@@

@@

XXXMore TextXXX from Amanda & AuroraXXX
Proteins derived from clade HLXXX ecotypes of *P. marinus* were detected in OMZ at depths up to 200 meters,  with O~2~ of 15 µM. The extent to which [O~2~] defines the niches occupied by different *P. marinus* ecotypes, as compared to potentially covarying environmental variables like photoperiod, light spectrum, and light level, is poorly described.

@@

NATLProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (clade HLI), SS120 (clade LLIII) and MIT9313 (clade LLIV).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

### MED4 clade HLI {.unnumbered}

Under 250 µM O~2~ the growth rates of *P. marinus* MED4, clade HLI, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), across all spectral wavebands. No growth was observed under any imposed conditions under a 4 h photoperiod. The maximum growth rate (µ~max~) was `r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 180 µE blue light and 16 h photoperiod.

@@

The GAM model in Figure \@ref(fig:MED4GAM) summarizes MED4 growth responses to red (A,B) or blue (C,D) peak PAR and photoperiod across 2 imposed oxygen concentrations.  Under 250 µM O~2~ MED4 achieved fastest growth rates above peak blue light of ~180 µmol photons m^-2^ s^-1^, and the longest photoperiod of 16, indicated by the 0.64 d^-1^ contour line representing the 90^th^ percentile of maximum achieved growth rate (Figure \@ref(fig:MED4GAM)C).  Growth decreased with decreasing photoperiod and decreasing peak PAR. Under red light growth was generally slower but the pattern of growth responses to photoperiod and PAR was similar (Figure \@ref(fig:MED4GAM)A).  Note the exclusion of MED4 from growth under 4 h photoperiod under both red and blue light (Figure \@ref(fig:MED4GAM)). Under 25 µM O~2~ MED4 showed similar growth responses, but was excluded from both 4 and 8 h photoperiods.  MED4 did not grow under 2.5 µM O~2~, so no GAM model was run. 
Considering the range of PAR levels, and spectral bands that MED4 can utilize, MED4 can inhabit not just shallow depths, where light levels are high, but also deeper regions, characterized by a lower level of blue light, subject to the limitation of a photoperiod of more than 4 h, even after depth attenuation of light. The photoregimes of winter temperate zones, due to shorter photoperiods, exclude MED4 from growth at any depth, however temperate photoperiods and light levels for the remainder of the year are potentially adequate to support MED4 growth, if water temperatures warm into the clade HLI tolerance range.

@@

### SS120 clade LLIII {.unnumbered}

Under 250 µM O~2~, growth rates of *P. marinus* SS120, clade LLIII, increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and across all spectral wavebands (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^ but showed growth inhibition at 16 h red light photoperiod.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white LED, red or blue light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

@@

### MIT9313 clade LLIV {.unnumbered}

Under 250 µM O~2~ growth rates of *P. marinus* MIT9313, clade LLIV, increased with longer
photoperiods, under low 30 µmol photons m^-2^ s^-1^ peak PAR, (Figure \@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiods. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while MIT9313 showed only marginal growth under white LED and red light at 180 µmol photons m^-2^ s^-1^ peak PAR, under the 8 h photoperiod, consistent with Moore *et al*. [@moorePhotophysiologyMarineCyanobacterium1999]. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 30 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod.

@@

### MED4 clade HLI {.unnumbered}

@@

### SS120 clade LLIII {.unnumbered}

@@

### MIT9313 clade LLIV {.unnumbered}

@@

Under full atmospheric [O~2~] and blue light, LL clades of *Prochlorococcus* are restricted to growth under low light, in part because they suffer photoinhibition of Photosystem II (PSII) through several paths, including direct absorbance of UV or blue light, in parallel with generation of Reactive Oxygen Species (ROS) if the electron flow is slowed [@aroPhotoinhibitionPhotosystemII1993], producing damaging singlet oxygen (^1^O~2~) [@aroPhotoinhibitionPhotosystemII1993; @soitamoPhotoinhibitionMarinePicocyanobacteria2017; @murphyPhotoinactivationPhotosystemII2017; @hakalaEvidenceRoleOxygenevolving2005]. The presence of NATL2A, a clade LLI, has been found near the ocean surface during deep ocean mixing [@malmstromTemporalDynamicsProchlorococcus2010]. Malmstrom *et al*. attributes NATL2a tolerance to short exposures of high light to the presence of the genes encoding photolyase, a gene found in HL clades which, in the presence of blue light, is responsible for repairing DNA damaged by UV light [@sancarStructureFunctionDNA2003]. Repair of photoinactivated PSII relies on the removal of damaged PsbA [@mannInvolvementFtsHHomologue2000; @adamFtsHProteasesChloroplasts2005], followed by reassembly with newly synthesized PsbA [@nixonRecentAdvancesUnderstanding2010]. Degradation of PsbA is a rate-limiting step in recovery from photoinhibition [@kanervoD1ProteinDegradation1993], mediated largely by a heterohexamer of (FtsH12)~3~, a membrane-bound (Sacharz et al., 2015)(Zak et al., 2001) metalloprotease  (Chiba et al., 2002) (Yoshioka-Nishimura and Yamamoto, 2014).  [@adamFtsHProteasesChloroplasts2005; @boehmSubunitOrganizationSynechocystis2012a; @nixonRecentAdvancesUnderstanding2010; @komendaExposedNTerminalTail2007] (Pisareva et al., 2007; Sacharz et al., 2015),  @mannInvolvementFtsHHomologue2000, (AMANDA REF from ALGATECH).

*Prochlorococcus* genomes encode 4 FtsH proteins, henceforth referred to as FtsH1-4, homologs to the characterized FtsH isoforms of the model freshwater cyanobacterium *Synechocystis* FtsH, and with presumably parallel functions (Table \@ref(tab:FtsHTable)). Upon a shift to higher light HLI MED4 upregulated expression of FtsH1 and FtsH2 [@bonisteelStrainSpecificDifferences2018], homologs to the *Synechocystis* slr0228 and slr1604, implicated in PSII repair. In contrast, representative LLIV strain MIT9313 showed less overall expression of the FtsH proteases, and thus has fewer FtsH serving each photosystem.  Furthermore, MIT9313 expressed primarily FtsH3, homologous to *Synechocystis* slr1463 involved in PSI biogenesis, and  FtsH expression did not increase in response to light stress in MIT9313. Through adaptation to steady low light, clade LLIV *Prochlorococcus* instead allocate resources to processes other than dynamic regulation of PSII repair.

@@

Ocean detections of proteins mediating protein metabolism support this interpretation of distinct FtsH function across clades of *P. marinus*. Ribosome proteins from clade HLI MED4, clade LLIII SS120 and clade LLIV MIT9313 show generally similar patterns vs. [O~2~] and depth, a proxy for peak PAR.  FtsH3, inferred to mediate PSI assembly, likewise shows a similar pattern between MED4 and MIT9313.  But only MED4 shows detected presence of the FtsH1 & FtsH2 isoforms inferred to mediate PSII repair, and then only in near-surface samples subject to higher light.  Furthermore, even though MIT9313 grows (Figure \@ref(fig:MIT9313GAM)), and is detected in the ocean at low [O~2~] (Figure \@ref(fig:NATLProchloroPhotosynthDepthO2)), no FtsH from MIT9313 is detected at low [O~2~], suggesting limited requirement for protein turnover under low [O~2~].

@@

NATLProchloroProteinMetabDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* protein metabolism complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (clade HLI), SS120 (clade LLIII) and MIT9313 (clade LLIV).",
                       "Columns show detections of proteins annotated as FtsH Protease Complexes (FtsH1, FtsH2, FtsH3) or the Ribosome).",
                       "Culture growth experimental conditions indicated by horizontal grey lines for depths approximating Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^) and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

@@

*Prochlorococcus marinus* SS120, a LLIII clade representative, showed an interactive inhibition of growth by oxygen and cumulative diel PUR, with a higher tolerance for higher cumulative diel PUR under 25 µM O~2~, compared to 250 µM O~2~. Thus, SS120 can exploit habitats with O~2~ levels spanning 25 µM O~2~ to 250 µM O~2~, including high PAR environments within OMZ.  SS120 is likely excluded from higher [O~2~] and higher PAR because of genomic limitations on capacity for DNA repair (Figure \@ref(fig:DNARepairFig)), and possibly by limited capacity for synthesis of reactive oxygen quenchers (Figure  \@ref(fig:ProchlorococcusEnzymeKms)).  Our growth results are supported by @lavinNovelLineagesProchlorococcus2010 who found evidence of LLII/III
and LLIV ecotypes, using terminal restriction fragment length polymorphism analyses, at depths above 40 m, where light levels are
higher, within OMZ, and by Figure \@ref(fig:NATLProchloroPhotosynthDepthO2) showing PSII protein subunits annotated as derived from SS120 at all depths ranging
from 20 to 200 m and all [O~2~] in an OMZ of the tropical North Pacific Ocean. SS120, grew under photoperiods longer than 4 h and showed increasing growth rate with increasing photoperiods. Therefore, SS120 can potentially exploit the longer summer photoperiods in temperate zones but again cannot sustain growth in winter through its requirement for more than 4 h of light per day, even if water temperature warms into the clade LLIII tolerance range.
